---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->

`Q7` enables _Compositional Object Oriented Programming_, a fluid, powerful and _postmodern_ paradigm, featuring:

#### Smart Objects
  - functions within an object know:
    - Where am I? What are my neighbors? 
  - extensible 
    - Make variants of an object

#### No Magic
  - Mechanism decomposes into basic R constructs 
    - A _type_ is a function
    - A _feature_ is a function
    - An _instance_, created by _type_ and _feature_, is an environment
  - Same great R syntax & semantics
    - Perform any action on or within an object 
    - Native scoping rules, almost no NSE
  
#### Compositional Manner of Construction
  - Freely add, change or delete members, ad or post hoc
  - Focuses on "has a", rather than than "is a" relationships 
  - Objects can contain other objects
  
#### Unlocked Objects
  - Object instances are unlocked environments
  - Feel free to shoot your feet, if so desired
  - Want safety? Lock'em yerself


`Q7` users can leave behind the grand narrative of the OOP orthodoxy, and exploit the benefits of objects simply as a unit of code, and as an instrument for namespace resolution.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```
## Examples

Let's make a `String` type and re-pack some base R functions into it, giving the object a distinct API, which some users may find familar with. 

```{r}
require(Q7)
String <- type(function(string){
  if (length(string) != 1 || !is.character(string)) { 
    # we only want this object to hold a 1-length character vector
    stop("string must be a vector of length 1")
  }
  .my$string <- string
  
  .my$charAt <- function(index){
    unlist(strsplit(.my$string, ""))[index + 1]
  }
  
  .my$concat <- function(str){
    stopifnot(inherits(str, 
                       "String"))
    
    .my$string <- paste0(.my$string, 
                         str$string)
    .my
  }
  
  .my$length <- function(){
    nchar(.my$string)
  }
  
  .my$isEmpty <- function(){
    nchar(.my$string) == 0
  }
  
  .my$matches <- function(pattern){
    .my$string <- grepl(pattern, .my$string)
    .my
  }
  
  .my$replaceFirst <- function(regex, replacement){
    .my$string <- sub(regex, 
                      replacement, 
                      .my$string)
    .my
  }
}, s3 = "String")

print.String <- function(String){
  if (inherits(String, "Q7instance")) {
    print(String$string)
  } else if (inherits(String, "Q7type")) {
    print(String)
  }
}
```

We have defined a type named _String_, which will produce instances with a S3 class _String_, as well as a S3 method for `print()`. The `$string` member is the vector that holds the data; All other members are bound functions (methods) of the object. They understand that they live inside the `String` object. The `print.String()` function lives in the global scope, and simply reaches into the `String` object and print the `$string$` vector.

`.my` object is a special object that refers to the instance itself. Although it can be safely excluded in most cases, with some precaution in place, the use of `.my` is highly recommended to avoid scope leak.

```{r}
motto <- String("Carpe")$concat(String(" "))$concat(String("Diem"))

motto$length()
motto$isEmpty()
motto$charAt(5)
motto$replaceFirst("\\s", "-")$charAt(5)
motto
```


### API
- `type()`
  - Defines a _type_. (like _class_)
  - Takes a function or expression as constructor
  - When invoked, the constructor's closure becomes an _instance_, which is an environment
    - Contains every binding inside the closure, except for the arguments
    - The arguments are not accessible outside of the object, making them private
    - Also contains `.my`, which refers to the instance itself
- `feature()`
  - Defines a _feature_
  - Takes an expression
  - Appends the expression to the object
    - Ad hoc: A _feature_ can be implemented on a _type_
    - Post hoc: Can also be implemented on an _instance_ 
- `implement()`
  - Takes
    - object, a _type_ or _instance_
    - any expression (including _features_, but more importantly, an arbitrary expression)
  - Appends the expresseion to the object

  

```{r include=FALSE}
require(Q7)
```

Make a type:
```{r}
TypeOne <- type(function(arg1, arg2){
  
  
})
```
`type()` takes a function, which is used to construct an instance. 

Everything defined within the function's closure becomes members of the object.
Arguments supplied to the function are accesible to the closure, but not become members of the object themselves. 
```{r}
type_one <- TypeOne(1, 2)
type_one$...
```

The object can be modified post-hoc. 

```{r}
type_one %>% implement({
  
  
})
```

The features implemented can be packaged with `feature()`. 
```{r}
hasFeatureOne <- feature({
  x <- 1
  y <- function(n){
    x + n
  }
})
```

```{r}
hasFeatureTwo <- feature({
  x <- 2 # Overrides x from hasFeatureOne
  old_y <- y
  y <- function(n){
    old_y(n) + 1
  }
})

```

#### Private members
```{r}

Counter <- type(function(count = 0){
    add_one <- function(){
      count <<- count + 1
    }
    
    get_count <- function(){
      count
    }
})

```

```{r}
counter <- Counter()
ls(counter)
counter$get_count()
counter$add_one()
counter$add_one()
counter$get_count()
```

##### Get Access to Private Environment
The private environment is upstream to (parent of) the public environment. The following code allows direct outside access to the `count` object.
```{r}
exposePrivateEnv <- feature({
  .private <- parent.env(.my)
})

counter %>% exposePrivateEnv()
counter$.private
counter$.private$count
```


#### Parameterized features?

_feature_ is subordinate to and dependent on _type_. It is encouraged to put all data members in a _type_ definition, while _feature_ mainly contain functions. If you feel significant need to parameterize a feature, think if it's better to create a nested object or to formally extend a type. You can always re-define something in a feature post hoc. 

This will be implemented in the future.

```{r}
Word <- type(function(word){})
hasRepeat <- feature({
  N_TIMES <- 2
  repeat_word <- function(){
    cat(rep(word, N_TIMES))
  }
})
```

```{r}
apple <- Word("apple") %>% hasRepeat()
apple$repeat_word()
```
```{r}
pear <- Word("pear") %>% 
  implement({
    hasRepeat()
    N_TIMES <- 5
  })
pear$repeat_word()
```

```{r}
repeatWordNTimes <- function(word, times){
  localize(Word)(word) %>% 
    hasRepeat() %>% 
    implement({
      N_TIMES <- times
    })
}

orange <- repeatWordNTimes("orange", 7)
orange$repeat_word()
```

