---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->

`Q7` provides _Freestyle Object Oriented_ Programming, a fluid, powerful and _postmodern_ paradigm that borrows from Go and Rust, featuring:

```{r}
library(Q7)
```


Make a type:
```{r}
typeOne <- type(function(arg1, arg2){
  
  
})
```
`type()` takes a function, which is used to construct an instance. 

Everything defined within the function's closure becomes elements of the object.
Arguments supplied to the function are accesible to the closure, but not become elements of the object themselves. 
```{r}
type_one <- typeOne(1, 2)
type_one$...
```

The object can be modified post-hoc. 

```{r}
type_one %>% implement({
  
  
})
```

The features implemented can be packaged with `feature()`. 
```{r}

```

```{r}

```




#### Smart Objects
  - self-aware 
    - Knows about itself
  - active
    - Bind functions within
  - extensible
    - Make variants of an object (class and instance)

#### No Magic
  - Mechanism decomposes into basic R constructs 
    - A type is a function
    - A feature is a function
    - An instance is an environment
  - Same great R syntax & semantics
    - Straightforwardly perform any action on or within an object 
    - Follows native scoping rules, almost no NSE
  
#### Compositional
  - ...not quite hereditary
  - Freely add, change or delete elements, ad/post hoc
  - Focuses on "has_a", rather than than "is_a" relationships
  - Objects can contain other objects (what is this called, Reference Semantics?)
  
#### Unlocked
  - Instances are unlocked environments
  - No one stops you from shooting your feet
  - Want safety? Lock'em yourself

### Interface

- `type()`
  - Defines a _type_. (like _class_)
  - Takes a function
  - Returns the same function, plus some other code
  - When invoked, the function's closure becomes an _instance_, which is an environment
    - Contains every binding inside the closure, except for the arguments
    - The arguments are not accessible outside of the object, making them private
    - Also contains `.my`, which refers to the instance itself
- `feature()`
  - Defines a _feature_
  - Takes any expression
  - Appends the expression to the object
    - Ad hoc: A _feature_ can be implemented on a _type_
    - Post hoc: Can also be implemented on an _instance_ 
- `implement()`
  - Takes
    - object, a _type_ or _instance_
    - any expression (including _features_, but more importantly, any arbitrary expression)
  - Appends the expresseion to the object

`Q7` users should leave behind the grand narrative of classical OOP orthodoxy, and exploit the benefits of objects as a unit of code, and an instrument for namespace resolution.
  
### versus R6
Q7: implicit definition private elements
R6: explicit definition of private 

Q7: can add or change features on-the-fly
R6: Must unlock object first; No apparent equivalent






## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```
## Examples

Walk through the following comment-free examples and see if you can figure out how `Q7` works.

```{r include=FALSE}
require(Q7)
```


### Dogs & Humans


```{r}
require(Q7)

Dog <- type(function(name, breed){
    self_intro <- function() {
        paste("My name is", name, "and I'm a", breed)
    }
    fav_food <- NULL
    fav_food2 <- NULL
    set_fav_foods <- function(food, food2) {
        .my$fav_food <- food
        fav_food2 <<- food2
    }
    collar <-  type(function() {
        color <- "red"
        buckle <- type(function(material = "gold"){
            material <- material
        })()
        brand <- "unknown"
    })()
})

Dog


my_dog <- Dog("Captain Cook", "Boston Terrier")
my_dog$self_intro()
my_dog$fav_food
my_dog$set_fav_foods("Sausage", "Bacon")
my_dog$fav_food
my_dog$fav_food2

new_dog <- clone(my_dog)
new_dog
my_dog

new_dog$name <- "Snowy"
new_dog$breed <- "Westie"
new_dog$self_intro()
my_dog$self_intro()

my_dog$collar
new_dog$collar$color <- "black"

my_dog$collar$color
new_dog$collar$color

identical(my_dog$collar$buckle, new_dog$collar$buckle)

my_dog %>% implement({
    owner <- NULL
    come_to_owner <- function(){
        paste(name, "runs toward", owner, 
              "in a collar that is", collar$color)
    }
})

my_dog$owner <- "Jack"
my_dog$come_to_owner()

ur_dog <- Dog("Fifi", "Bulldog")
ur_dog$collar$color


```



```{r}
Dog <- type(function(name, breed){
    say <- function(greeting = "Woof!"){
        cat(paste0(greeting, 
                   " I am ", name, ", a ", breed, 
                   ".\n"))
    }    
})
```


```{r}
walter <- Dog("Walter", "Husky")
ls(walter, all.names = TRUE)
```


```{r}
walter$say()
```

```{r}
max <- walter %>% clone()
max$name <- "Max"
max$say("Wussup Dawg!")
```

```{r}
max %>% 
  implement({
    treats_eaten <- 0
    eat_treat <- function(n = 1){
      cat(paste(name, "eats", n, "treat(s).\n"))
      treats_eaten <<- treats_eaten + n
    }
  }) %>% 
  implement({
    is_satisfied <- function(){
      treats_eaten > 5
    }
  })
```


```{r}
max$eat_treat()
max$is_satisfied()
max$eat_treat(2)
max$is_satisfied()
max$eat_treat(3)
max$is_satisfied()
```

```{r}
max$treats_eaten
```

```{r}
isAnimal <- feature({
    mortal <- TRUE
    eat <- function() paste(.my$name, "eats.")
    poop <- function() paste(name, "poops")
})

max %>% 
  isAnimal()

max$eat()
max$poop()
```

```{r}
Person <- type(function(name, job) {
  description <- function(){
    paste(name, "works as a(n)", job)
  }
}) %>% isAnimal()
archie <- Person("Archie", "Analyst")
```


```{r}
archie
archie$description()
archie$poop()
```

```{r}
hasCollar <- feature({
  Collar <- type(function(material, color){
    description <- function() {
      paste("is made of", material, "and in", color)
    }
  })
  
  take_for_a_walk <- function(){
    cat(name, "wears a collar that", collar$description(), "\n")
    cat("We're gonna go out for a walk!")
  }
})
```

```{r}
walter %>%
  implement({
    hasCollar()
    collar <- Collar("metal", "red")
    rm(Collar)
  })
```


```{r}
walter$take_for_a_walk()
```

### Formally Extend a Type
```{r}
Type1 <- type(function(arg1){
    val1 <- arg1
    get_val1 <- function(){
        val1
    }
}, "Type1")

Type2 <- type(function(arg1, arg2){
    extend(Type1)(arg1)
    val2 <- arg2
    get_val2 <- function(){
        val2
    }
}, "Type2")
```


```{r}
type2 <- Type2("one", "two")
type2$val1
type2$val2
type2$get_val1()
type2$get_val2()
```


### Overtime
```{r}
Employee <- type(function(weekly_hours){}, "Employee")
john <- Employee(45)
```

```{r}
Manager <- type(function(weekly_hours){
  extend(Employee)(weekly_hours)
  is_manager <- TRUE
}, "Manager")

mike <- Manager(45)
```


```{r}
hasOvertime <- feature_generic("hasOvertime")
  
hasOvertime.Employee <- feature({
  is_overtime <- function() weekly_hours > 40
})
hasOvertime.Manager <- feature({
  .my$is_overtime <- function() FALSE
})
```


```{r}
john %>% hasOvertime()
john$is_overtime()
```

```{r}
mike %>% hasOvertime()
mike$is_overtime()
```


```{r}
Boss <- 
  type(function(weekly_hours){
    weekly_hours <- weekly_hours
  }) %>% 
  hasOvertime.Manager()
jill <- Boss(80)
jill$is_overtime()
```

### List-to-Instance
```{r}
my_data <- list(a = 1, 
                add_to_a = function(value){
                  .my$a <- a + value
                })

my_data_obj <- list2inst(my_data)

my_data_obj$a
my_data_obj$add_to_a(20)
my_data_obj$a
```

### Grade School Geometry
```{r}
Circle <- 
    type(function(radius){}, 
         "Circle")
    
Square <- 
    type(function(side){}, 
         "Square")

hasArea <- feature_generic("hasArea")

hasArea.Square <- 
    feature({
        area <- function(){
            .my$side ^ 2
        }
    })

hasArea.Circle <- 
    feature({
        area <- function(){
            .my$radius^2 * pi
        }
    })

circle <- Circle(1) %>% hasArea()
circle$area()

square <- Square(1) %>% hasArea()
square$area()


hasArea.EquilateralTriangle <- feature({
    area <- function(){
        (side^2 * sqrt(3)) / 4
    }
})

EquilateralTriangle <- 
    type(function(side){}, 
         "EquilateralTriangle") %>%
    hasArea()

equilateral_triangle <- EquilateralTriangle(1)
equilateral_triangle$area()

```


### Flying Rat
```{r}
Rat <- type(function(){}, "Rat")
hasWing <- feature({
  can_fly <- TRUE
})
Pigeon <- Rat %>% hasWing()
pigeon <- Pigeon()
pigeon$can_fly
```


### Locked 

```{r}
isLocked <- feature({
    lockEnvironment(.my, bindings = TRUE)
})

Test <- type(function(){
    a <- 1
}) %>% isLocked()

test <- Test()
try(test$a <- 666)
try(test$b <- 666)
try({
  test %>% 
    implement({
      a <- 666
    })
})
```
### State Machine
```{r}
State <- type(function(){
    name <- "DEFAULT"
    cat("Processing Current State...\n")
    print_current_state <- function(){
        cat(paste("Current State:", name, "\n"))
    }
})

LockedState <- State %>%
    implement({
        name <- "Locked"
        print_current_state()
        on_event <- function(event) {
            if (event == "8888") {
                return(UnlockedState())
            } else {
                cat("Wrong Password.\n")
                return(.my)
            }
        }
    })

UnlockedState <- State %>%
    implement({
        name <- "Unlocked"
        print_current_state()
        on_event <- function(event) {
            if (event == "lock") {
                return(LockedState())
            } else {
                cat("Invalid Operation. \n")
                return(.my)
            }
        }
    })

SimpleDevice <- type(function(){
    state <- LockedState()
    on_event <- function(event){
        .my$state <- state$on_event(event)
    }
})
```


```{r}
device <- SimpleDevice()

device$on_event("0000")
device$on_event("8888")
device$on_event("do something")
device$on_event("lock")
```

Private Elements
```{r}

Counter <- type(function(count = 0){
    add_one <- function(){
      count <<- count + 1
    }
    
    get_count <- function(){
      count
    }
})

```

```{r}
counter <- Counter()
ls(counter)
counter$get_count()
counter$add_one()
counter$add_one()
counter$get_count()
```


```{r}
R6Example <- R6::R6Class("R6Example", 
                         public = list(
                             a = 1, 
                             b = 2, 
                             f1 = function(){
                                 self$c <- self$a + self$b
                             }, 
                             c = NULL,
                             f2 = function(){
                                 private$d <- self$a + self$b
                             }, 
                             f3 = function(){
                                 private$d
                             }, 
                             initialize = function(){
                                 cat("initializing...")
                             }
                         ), 
                         private = list(
                             d = NULL
                         )
)

r6 <- R6Example$new()




require(foo)
Q7Example <- type(function(var4){
    cat("initializing...")
    var1 <- 1
    var2 <- 2
    fn1 <- function(){
        var3 <<-  var1 + var2
    }
    var3 <- NULL
    fn2 <- function(){
        var4 <<- var1 + var2
    }
    fn3 <- function(){
        var4
    }
})


Q7 <- Q7Example()
Q7$fn2()
Q7$fn3()

```

