---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# foo

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->

Welcome! `foo` provides _Freestyle Object Oriented_ Programming, featuring:

- Native R Flavor
  - Everything is familiar
  - Use `<-` to assign value to bindings
  
- No Inheritance
  - No lineage to trace back to
  - No methods you don't need
  - No mess
  
- Compostable
  - _Reference Semantics_
  - instance can contain other instances
  - post-hoc `implement()` to add more custom features

- Unlocked environment/binding
  - Each object is an environment
  - use `.my` to refer to self (optional)
  - Deep copy by default
  - add, remove, change items in an object freely as you please
  - want immutability? don't write the code that changes things!

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/foo")
```
## Example

Walk through the following example and see if you can figure out how `foo` works.

```{r include=FALSE}
require(magrittr)
require(foo)
```


```{r}
Dog <- type(function(name, breed){
    say <- function(greeting = "Woof!"){
        cat(paste0(greeting, 
                   " I am ", name, ", a ", breed, 
                   ".\n"))
    }    
})
```


```{r}
walter <- Dog("Walter", "Husky")
ls(walter, all.names = TRUE)
```


```{r}
walter$say()
```

```{r}
max <- walter %>% clone()
max$name <- "Max"
max$say("Wussup Dawg!")
```

```{r}
max %>% 
  implement({
    treats_eaten <- 0
    eat_treat <- function(n = 1){
      cat(paste(name, "eats", n, "treat(s).\n"))
      treats_eaten <<- treats_eaten + n
    }
  }) %>% 
  implement({
    is_satisfied <- function(){
      treats_eaten > 5
    }
  })
```


```{r}
max$eat_treat()
max$is_satisfied()
max$eat_treat(2)
max$is_satisfied()
max$eat_treat(3)
max$is_satisfied()
```

```{r}
max$treats_eaten
```

```{r}
isAnimal <- feature({
    mortal <- TRUE
    eat <- function() paste(.my$name, "eats.")
    poop <- function() paste(name, "poops")
})

max %>% 
  isAnimal()

max$eat()
max$poop()
```

```{r}
Person <- type(function(name, job) {
  description <- function(){
    paste(name, "works as a(n)", job)
  }
  isAnimal()
})
archie <- Person("Archie", "Analyst")
```


```{r}
archie$description()
archie$poop()
```

```{r}
hasCollar <- feature({
  Collar <- type(function(material, color){
    description <- function() {
      paste("is made of", material, "and in", color)
    }
  })
  
  take_for_a_walk <- function(){
    cat(name, "wears a collar that", collar$description(), "\n")
    cat("We're gonna go out for a walk!")
  }
})
```

```{r}
walter %>%
  implement({
    hasCollar()
    collar <- Collar("metal", "red")
    rm(Collar)
  })
```


```{r}
walter$take_for_a_walk()
```

```{r}
Employee <- type(function(weekly_hours){NULL}, "Employee")
john <- Employee(45)
```

```{r}
Manager <- type(function(weekly_hours){NULL}, "Manager")
mike <- Manager(45)
```


```{r}
hasOvertime <- feature_generic("hasOvertime")
  
hasOvertime.Employee <- feature({
  is_overtime <- function() weekly_hours > 40
})
hasOvertime.Manager <- feature({
  is_overtime <- function() FALSE
})
```


```{r}
john %>% 
  hasOvertime()
john$is_overtime()
```

```{r}
mike %>% 
  implement(hasOvertime())
mike$is_overtime()
```


```{r}
Boss <- 
  type(function(weekly_hours){NULL}) %>% 
  hasOvertime.Manager()
jill <- Boss(80)
jill$is_overtime()
```

```{r}
my_data <- list(a = 1, 
                add_to_a = function(value){
                  .my$a <- a + value
                })

my_data_obj <- list2inst(my_data)

my_data_obj$a
my_data_obj$add_to_a(20)
my_data_obj$a
```

## Another Example
```{r}
require(foo)

Circle <- 
    type(function(radius){NULL}, 
         "Circle")
    
Square <- 
    type(function(side){NULL}, 
         "Square")

hasArea <- feature_generic("hasArea")

hasArea.Square <- 
    feature({
        area <- function(){
            .my$side ^ 2
        }
    })

hasArea.Circle <- 
    feature({
        area <- function(){
            .my$radius^2 * pi
        }
    })

circle_1 <- Circle(1)
circle_1 %>% hasArea()
circle_1$area()

square_1 <- Square(1)
square_1 %>% hasArea()
square_1$area()


hasArea.EquilateralTriangle <- feature({
    area <- function(){
        (side^2 * sqrt(3)) / 4
    }
})

EquilateralTriangle <- 
    type(function(side){NULL}, 
         "EquilateralTriangle") %>%
    hasArea()

equilateral_triangle_1 <- EquilateralTriangle(1)
equilateral_triangle_1$area()

```

```{r}
Rat <- type(function(){NULL}, "Rat")
hasWing <- feature({
  can_fly <- TRUE
})
Bat <- Rat %>% 
  hasWing()
bat <- Bat()
bat$can_fly
```


TODO: what if features have same bidings? what to do then? 
 - evaluate each feature to their own environment, each has the parent of .my, or inside .my
 - let the user pick and choose what bindings to _import_
 - prefix bidings with class name

