---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->

Q7 enables a postmodern flavor of object-oriented programming (OOP), a flexible and powerful paradigm. Q7 features: 

- Compositional manner of construction with `type()`, `feature()` and `implement()`
- Has `public`, `private`, `final` and `active` binding modifiers
- 

`Q7` users can leave behind the grand narrative of the OOP orthodoxy, and enjoy 



#### Smart Objects
  - functions within an object know:
    - Where am I? What are my neighbors? 
  - extensible 
    - Make variants of an object

#### No Magic
  - Mechanism decomposes into basic R constructs 
    - A _type_ is a function
    - A _feature_ is a function
    - An _instance_, created by _type_, is an environment
  - Same great R syntax & semantics
    - Perform any action on or within an object 
    - Normal scoping rules
  
#### Compositional Manner of Construction
  - Freely add, change or delete members, ad or post hoc
  - Focuses on "has a", rather than than "is a" relationships 
  - Objects can contain other objects
  
## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```

### Example

To make an object: 
```{r}
require(Q7)
Adder <- type(function(num1, num2){
  
  
})
```
