---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# foo

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->


`foo` provides _Freestyle Object Oriented_ Programming, a fluid & powerful paradigm that has many creative uses, featuring:

- Smart Objects: 
  - self-aware 
    - Knows about itself
  - active
    - Stores & invokes functions within
  - extensible
    - Make variants of an object

- No Magic
  - Mechanism decomposes into basic R constructs 
    - A type is a function
    - A feature is a function
    - An instance is an environment
  - Same great R syntax & semantics
    - Straightforwardly perform any action on or within an object 
    - Follows native lexical scoping rules, almost no NSE
  
- Compositional
  - ...not hereditary
  - Freely add, change or delete elements, ad/post hoc
  - Focuses on "has-a", rather than than "is_a" relationships
  - Objects can contain other objects (Reference Semantics?)
  
- Mutable 
  - Instances are unlocked environments
  - Easy run time debugging
  - No one stops you from shooting your feet
  - Want safety? Lock'em yourself

Interface

- `type()`
  - Defines a _type_. (like _class_)
  - Takes a function
  - Returns the same function, plus some other code
  - When invoked, the function's closure becomes an _instance_, which is an environment
    - Contains every binding inside the closure, including arguments
    - Also contains `.my`, which refers to the instance itself
- `feature()`
  - Defines a _feature_
  - Takes any expression
  - Appends the expression to the object
    - Ad hoc: A _feature_ can be implemented on a _type_
    - Post hoc: Can also be implemented on an _instance_ 
- `implement()`
  - Takes
    - object, a _type_ or _instance_
    - any expression (including _features_, but more importantly, any arbitrary expression)
  - Returns what was passed in
  - Appends the expresseion to the object

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/foo")
```
## Example

Walk through the following examples and see if you can figure out how `foo` works.

```{r include=FALSE}
require(foo)
```


## Dogs & Humans
```{r}
Dog <- type(function(name, breed){
    say <- function(greeting = "Woof!"){
        cat(paste0(greeting, 
                   " I am ", name, ", a ", breed, 
                   ".\n"))
    }    
})
```


```{r}
walter <- Dog("Walter", "Husky")
ls(walter, all.names = TRUE)
```


```{r}
walter$say()
```

```{r}
max <- walter %>% clone()
max$name <- "Max"
max$say("Wussup Dawg!")
```

```{r}
max %>% 
  implement({
    treats_eaten <- 0
    eat_treat <- function(n = 1){
      cat(paste(name, "eats", n, "treat(s).\n"))
      treats_eaten <<- treats_eaten + n
    }
  }) %>% 
  implement({
    is_satisfied <- function(){
      treats_eaten > 5
    }
  })
```


```{r}
max$eat_treat()
max$is_satisfied()
max$eat_treat(2)
max$is_satisfied()
max$eat_treat(3)
max$is_satisfied()
```

```{r}
max$treats_eaten
```

```{r}
isAnimal <- feature({
    mortal <- TRUE
    eat <- function() paste(.my$name, "eats.")
    poop <- function() paste(name, "poops")
})

max %>% 
  isAnimal()

max$eat()
max$poop()
```

```{r}
Person <- type(function(name, job) {
  description <- function(){
    paste(name, "works as a(n)", job)
  }
  isAnimal()
})
archie <- Person("Archie", "Analyst")
```


```{r}
archie$description()
archie$poop()
```

```{r}
hasCollar <- feature({
  Collar <- type(function(material, color){
    description <- function() {
      paste("is made of", material, "and in", color)
    }
  })
  
  take_for_a_walk <- function(){
    cat(name, "wears a collar that", collar$description(), "\n")
    cat("We're gonna go out for a walk!")
  }
})
```

```{r}
walter %>%
  implement({
    hasCollar()
    collar <- Collar("metal", "red")
    rm(Collar)
  })
```


```{r}
walter$take_for_a_walk()
```


## Overtime
```{r}
Employee <- type(function(weekly_hours){}, "Employee")
john <- Employee(45)
```

```{r}
Manager <- type(function(weekly_hours){}, "Manager")
mike <- Manager(45)
```


```{r}
hasOvertime <- feature_generic("hasOvertime")
  
hasOvertime.Employee <- feature({
  is_overtime <- function() weekly_hours > 40
})
hasOvertime.Manager <- feature({
  is_overtime <- function() FALSE
})
```


```{r}
john %>% 
  hasOvertime()
john$is_overtime()
```

```{r}
mike %>% 
  implement(hasOvertime())
mike$is_overtime()
```


```{r}
Boss <- 
  type(function(weekly_hours){}) %>% 
  hasOvertime.Manager()
jill <- Boss(80)
jill$is_overtime()
```

## List-to-Instance
```{r}
my_data <- list(a = 1, 
                add_to_a = function(value){
                  .my$a <- a + value
                })

my_data_obj <- list2inst(my_data)

my_data_obj$a
my_data_obj$add_to_a(20)
my_data_obj$a
```

## Grade School Geometry
```{r}
require(foo)

Circle <- 
    type(function(radius){}, 
         "Circle")
    
Square <- 
    type(function(side){}, 
         "Square")

hasArea <- feature_generic("hasArea")

hasArea.Square <- 
    feature({
        area <- function(){
            .my$side ^ 2
        }
    })

hasArea.Circle <- 
    feature({
        area <- function(){
            .my$radius^2 * pi
        }
    })

circle_1 <- Circle(1) %>% hasArea()
circle_1$area()

square_1 <- Square(1) %>% hasArea()
square_1$area()


hasArea.EquilateralTriangle <- feature({
    area <- function(){
        (side^2 * sqrt(3)) / 4
    }
})

EquilateralTriangle <- 
    type(function(side){}, 
         "EquilateralTriangle") %>%
    hasArea()

equilateral_triangle_1 <- EquilateralTriangle(1)
equilateral_triangle_1$area()

```


## Flying Rat
```{r}
Rat <- type(function(){}, "Rat")
hasWing <- feature({
  can_fly <- TRUE
})
Pigeon <- Rat %>% hasWing()
pigeon <- Pigeon()
pigeon$can_fly
```


## Locked 

```{r}
isLocked <- feature({
    lockEnvironment(.my, bindings = TRUE)
})

Test <- type(function(){
    a <- 1
}) %>% isLocked()

test <- Test()
try({
  test$a <- 666
  test$b <- 666
  
  test %>% implement({
    a <- 666
  })
})
```
## State Machine
```{r}
State <- type(function(){
    name <- "DEFAULT"
    cat("Processing Current State...\n")
    print_current_state <- function(){
        cat(paste("Current State:", name, "\n"))
    }
})

LockedState <- State %>%
    implement({
        name <- "Locked"
        print_current_state()
        on_event <- function(event) {
            if (event == "8888") {
                return(UnlockedState())
            } else {
                cat("Wrong Password.\n")
                return(.my)
            }
        }
    })

UnlockedState <- State %>%
    implement({
        name <- "Unlocked"
        print_current_state()
        on_event <- function(event) {
            if (event == "lock") {
                return(LockedState())
            } else {
                cat("Invalid Operation. \n")
                return(.my)
            }
        }
    })

SimpleDevice <- type(function(){
    state <- LockedState()
    on_event <- function(event){
        .my$state <- state$on_event(event)
    }
})

device <- SimpleDevice()

device$on_event("0000")
device$on_event("8888")
device$on_event("do something")
device$on_event("lock")
```


TODO: what if features have same bidings? what to do then? 

- evaluate each feature to their own environment, each has the parent of .my, or inside .my
- let the user pick and choose what bindings to _import_
- prefix bidings with type name

