---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# object

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

Welcome! `object` provides the following:

- Native R Style
  - Everything is familiar
  - Use `<-` to assign value to bindings as object's items
  
- No Inheritance
  - No mess
  - Easy to understand
  - Not unlike today's hottest languages: Go & Rust 
  
- Compostable
  - _Reference Semantics_
  - use `.my` to refer to self (optional)
  - Composition is better than Inheritance
  - Deep copy by default

- Unlocked environment/binding
  - Each object is an environment
  - add, remove, change items in an object freely as you please
  - want immutability? don't write the code that changes things!

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/object")
```
## Example

```{r}
dog <- object::type(function(name, breed){
    say <- function(greeting = "Woof!"){
        cat(paste0(greeting, 
                   " I am ", name, ", a ", breed, 
                   ".\n"))
    }    
})
```


```{r}
walter <- dog("Walter", "Husky")
ls(walter, all.names = TRUE)
```


```{r}
walter$say()
```

```{r}
max <- object::clone(walter)
max$name <- "Max"
max$say("Wussup Dawg!")
```

```{r}
max$.implement({
    treats_eaten <- 0
    eat_treat <- function(n = 1){
        cat(paste(name, "eats", n, "treat(s).\n"))
        treats_eaten <<- treats_eaten + n
    }
    
})
```


```{r}
max$eat_treat(2)
max$eat_treat(3)
```

```{r}
max$treats_eaten
```
```{r}
animal_traits <- object::feature({
    mortal <- TRUE
    eat <- function() paste(.my$name, "eats.")
    poop <- function() paste(name, "poops")
})

max$.implement({
    animal_traits()
})

max$eat()
max$poop()
```

```{r}
archie <- object::type(function(name = "Archie", race = "W") NULL )()
```

```{r}
Map(function(obj) obj$.implement(animal_traits()), 
    list(walter, archie))
walter$eat()
archie$poop()
```

