---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
[![Travis build status](https://travis-ci.org/iqis/q7.svg?branch=master)](https://travis-ci.org/iqis/q7)
<!-- badges: end -->

Q7 is a type system that enables a postmodern flavor of object-oriented programming (OOP), a simple and flexible paradigm, leaving behind the grand narrative of classical OOP. Q7 features: 

- `type()`, `feature()` and `implement()` to compose objects.
- `initialize()` and `finalize()` to run at an object's beginning and the end of life 
- `public`, `private`, `final`, `private_final` and `active` binding modifiers

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```
```{r}
require(Q7)
```

### Example
Make a Q7 object in 3 easy steps. 

1, Define an object type: 
```{r}
Adder <- type(function(num1, num2){
    add_nums <- function(){
        num1 + num2
    }
 })
```

2, Instantiate the object:
```{r}
myAdder <- Adder(1, 2)
```

3, Enjoy!
```{r}
myAdder$add_nums()
```

##### Smart Objects
  - Functions domestic to an object know:
    - Where am I? What are my neighbors? 
  - Extensible 
    - User can make variants of an object
  
##### Compositional Construction
  - Freely add, change or delete members, ad or post hoc
  - Focuses on _has a_, rather than than _is a_ relationships 
  - Objects can contain references to other objects

##### No Magic
  - Mechanism decomposes into basic R constructs 
    - A _type_ is a function
    - A _feature_ is a function
    - An _instance_, created by _type_, is an environment
  - Same great R syntax & semantics
    - Perform any action on or within an object 
    - Normal scoping rules

__Origin of the Name__ 
The package was named tounge-in-cheek `foo` for _Freestyle Object Orientation_, but the author soon realized the smart-a** name is going to backfire and cause real confusions. 
The latest OO system in R's ecosystem has been S3, S4, S5(RC) and R6. Desiring an air of seriousness, the author buttoned up and dialed the alphabet up and the numberal down: Q7, the next-in-line, yet with a distinct philosophy.
Trivia: Both R6 and Q7 are model names of motor vehicles.
