---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/iqis/q7.svg?branch=master)](https://travis-ci.org/iqis/q7)
<!-- badges: end -->

Q7 is a type system that enables a postmodern flavor of compositional object-oriented programming (OOP). 

It is simple, flexible and promotes healthy program design. No more family tree of classes! 
  
Q7 features: 

- `type()`, `feature()` and `implement()` to compose objects;
- For each object, `initialize()` and `finalize()` to run at its beginning and end of life;
- For each object, `public`, `private`, `final`, `private_final` and `active` binding modifiers to change the visibility and behavior of its members.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```
```{r}
require(Q7)
```

### Example
Make a Q7 object in 3 easy steps. 

1, Define an object type: 
```{r}
Adder <- type(function(num1, num2){
    add_nums <- function(){
        num1 + num2
    }
 })
```

2, Instantiate the object:
```{r}
myAdder <- Adder(1, 2)
```

3, Enjoy!
```{r}
myAdder$add_nums()
```

See vignettes for extending an object and other topics. 

### Features

##### Smart Objects
  - Contains reference to self: `.my`
  - Object methods know:
    - Which object do I belong to? 
    - Which are other members of the same object? 
  
##### Compositional Construction
  - Freely add, change or delete members, ad or post hoc, without subclassing
  - Focuses on _has-a_, rather than than _is-a_ relationships 
  - Objects can contain references to other objects

##### No Magic
  - All mechanisms are built from basic R constructs 
    - A _type_ is a function
    - A _feature_ is a function
    - An _instance_, created by _type_ and _feature_, is an environment
  - Same great R syntax & semantics
    - Perform any action on or within an object 
    - Normal scoping rules you would expect

__*Origin of the Name__ 

The package was named `foo` for *Freestyle Object
Orientation*, but the author soon realized the smart-a\*\* name is going to cause confusions. As it is more streamelined than R6, but represents progress in OO philosophy, the author dialed the alphabet down and the numberal up. 
