---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Q7

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
<!-- badges: end -->

Q7 enables a postmodern flavor of object-oriented programming (OOP), a simple and flexible paradigm, leaving behind the grand narrative of classical OOP. Q7 features: 

- Compose objects with `type()`, `feature()` and `implement()`
- `initialize()` and `finalize()` to run at an object's beginning and the end of life 
- `public`, `private`, `final` and `active` binding modifiers

## Installation

``` r
# install.packages("devtools")
devtools::install_github("iqis/Q7")
```
```{r}
require(Q7)
```

### Example
Make a Q7 object in 3 easy steps. 

1, Define an object type: 
```{r}
Adder <- type(function(num1, num2){
    add_nums <- function(){
        num1 + num2
    }
 })
```

2, Instantiate the object:
```{r}
myAdder <- Adder(1, 2)
```

3, Enjoy!
```{r}
myAdder$add_nums()
```

#### Smart Objects
  - Functions domestic to an object know:
    - Where am I? What are my neighbors? 
  - Extensible 
    - Make variants of an object
    
#### No Magic
  - Mechanism decomposes into basic R constructs 
    - A _type_ is a function
    - A _feature_ is a function
    - An _instance_, created by _type_, is an environment
  - Same great R syntax & semantics
    - Perform any action on or within an object 
    - Normal scoping rules
  
#### Compositional Manner of Construction
  - Freely add, change or delete members, ad or post hoc
  - Focuses on "has a", rather than than "is a" relationships 
  - Objects can contain other objects
