---
title: "OOP for Shiny"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{oop_shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  errors = TRUE
)
```

```{r setup}
library(Q7)
```

# OOP for Shiny

R is a functional programming language. Built with R, the web framework Shiny also employs this paradigm. 
This document shows one of many ways to employ OO patterns in a Shiny app with Q7 type system. 


Each module is an object that has the following members: 

 - `id`: constructor parameter, an aribitrary string to represent the whole module
 - `ns`: constructor parameter, a Shiny namespace function, defaults to depend on `id`, but can be overidden
 - `dev_mode`: constructor parameter, a simple boolean value to tell the module to turn some things on or off
 - `ui`: a _ui_ module fragment
 - `server`: definition and call to a _server_ module fragment
 - abstractions necessary to the above
 - `state`: an reference object made by `reactiveValues()` or `new.env()`, which esposes the what's meaningful to other modules.


```{r}
ShinyModule <- type(function(
  id = uuid::UUIDgenerate(FALSE), 
  ns = NS(id), 
  dev_mode = FALSE
){
  # The following are not strictly necessary, as they will be overidden,
  # but serve as good reminders
  ui <- function(){tagList(
    # ...
  )} 
  
  server <- function(){callModule(function(input, output, session){
    # ...
  }, id)}
  
  state <- reactiveValues()
})

LoadData <- ShinyModule %>% 
  implement({
    ui <- function(){
      
    }
    
    server
    
    
  })

PlotChart <- ShinyModule %>% 
  implement({
    ui
    
    server
    
    
  })

```

 
```{r eval=FALSE}
require(shiny)
require(Q7)

loadData <- LoadData()
plotChart <- PlotChart()

shinyApp(
  function(){fluidPage(sidebarLayout(  # UI
    loadData$ui(),
    plotChart$ui()
  ))},
  function(input, output, session){ # Server
    loadData$server()
    plotChart$server()
  }
)
```
 
Segregation of namespace and state. 

If you use `shinydashboard`, each is naturally a candidate for a top-level module. 


If you just need to define a input or output panel that has similarly organized and configured, you don't need a module.


