---
title: "Versus R6"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{versus_r6}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Q7)
```

### 

Q7: Compositonal: I put a plasma gun on my spaceship.
R6: Hereditary: I made my spaceship a special kind of spaceship that has a plasma gun.

Q7: Can add or change features on-the-fly
R6: Must unlock object first; No apparent equivalent mechanism

Q7: Free
R6: Safe

Q7: Succinct & sweet: one function defines & initializes the whole object
R6: Verbose & bitter: `public`, `private`, `active`, `$new()`, `$clone()`...

Q7: Implicitly defined private members
R6: Explicitly defined private members

Q7: Developed by one amateur
R6: Developed by a team of experts


```{r}
R6Example <- R6::R6Class("R6Example", 
                         public = list(
                           var1 = 1, 
                           var2 = NULL,
                           var3 = NULL,
                           f1 = function(){
                             self$var3 <- self$var1 + self$var2
                           }, 
                           f2 = function(){
                             private$var4 <- self$var1 + self$var2
                           }, 
                           f3 = function(){
                             private$var4
                           }, 
                           initialize = function(var2){
                             cat("initializing...")
                             self$var2 <- var2
                           }
                         ), 
                         private = list(
                           var4 = NULL
                         )
)

r6 <- R6Example$new(2)
r6$f1()
r6$var2
r6$var3
```


```{r}
require(Q7)
Q7Example <- type(function(var2, var4){
  cat("initializing...")
  var1 <- 1
  .my$var2 <- 2
  var3 <- NULL
  fn1 <- function(){
    var3 <<-  var1 + var2
  }
  fn2 <- function(){
    var4 <<- var1 + var2
  }
  fn3 <- function(){
    var4
  }
})


q7 <- Q7Example(2)
q7$fn1()
q7$fn2()
q7$fn3()
ls(q7)
```
